Fast solvers for the high-order FEM simplicial de Rham complex \thanks{Submitted to the editors July XX, 2024. \funding{ PDB and PEF were supported by EPSRC grant EP/W026260/1. } }

### 2024-06-17 paper/paper.tex
Some work on section 2

% Classic symbols
\newcommand{\reals}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\dx}{\,\d\mathbf{x}}
\newcommand{\md}{\,\d}
\newcommand{\bigo}[1]{\mathcal{O}(#1)}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: \#1}}

@@

\begin{figure}[htbp] 
\centering
\begin{tikzcd}
  \CG_p \arrow[r, "\grad"] & \Ned_p \arrow[r, "\curl"] & 
   \RT_p \arrow[r, "\div"] & \DG_{p-1} \\
  \Hgrad \arrow[r, "\grad"] \arrow[u] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
  \arrow[d] \arrow[u] & \Hdiv \arrow[r, "\div"] \arrow[u] \arrow[d] & \Ltwo \arrow[u] \arrow[d] \\
   \CG_p \arrow[r, "\grad"] & \NedTwo_{p-1} \arrow[r, "\curl"] & 
   \BDM_{p-2} \arrow[r, "\div"] & \DG_{p-3}
\end{tikzcd}
\caption{The $\Ltwo$-de Rham complex (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
\end{figure}

We follow the definition of interpolation degrees of freedom from
\cite{demkowicz00} to construct dual bases that promote orthogonality in the
$\Ltwo(\Khat)$ and $H(\mathrm{d}, \Khat)$ inner products.

We denote the space of polynomials with vanishing boundary trace as
$\mathbb{B}_{p}(\mathrm{d}, K) = [\P_{p}(K)]^d \cap H_0(\mathrm{d}, K)$.

@@

Let $\Delta^d$ be the equilateral $d$-simplex.
We define a basis for the dual of $\CG_p$ as 
point evaluations at $V \in \text{vertices}(\Delta^d)$,
\begin{equation}
   \ell^V(v) = v(\bx_V),
\end{equation}
and for each sub-entity $S \in \text{edges}(\Delta^d) \cup \text{faces}(\Delta^d) \cup \text{interior}(\Delta^d)$
we take integral moments of surface gradients, 
\begin{equation}
   \ell^S_j(v) = (\grad_S\phi^S_j, \grad_S v)_{S},
\end{equation}
where $\grad_S$ is the tangential gradient on $S$,
and $\{\phi^S_j\}$ is a basis for $\mathbb{B}_{p}(\grad, S)$, such that
\begin{equation} \label{eq:hgrad-fdm}
   (\grad_S\phi^S_j, \grad_S\phi^S_i)_{S} = \delta_{ij}, \quad
   (\phi^S_j, \phi^S_i)_{S} = \lambda_j\delta_{ij}.
\end{equation}
The eigenbases $\{\phi^S_j\}$ are numerically computed offline and only once 
on the reference interval, triangle, and tetrahedron.

@@

We define a basis for the dual of $V^1_{p, p_E}(K)$ as 
tangential moments along $E\in \text{edges}(\Delta^d)$,
\begin{equation}
   \ell^E_j(v) = (q_j, v\cdot \bt)_E, \quad q_j \in \P_{p_E}(E), 
\end{equation}
and for each sub-entity $S \in \text{faces}(\Delta^d) \cup \text{interior}(\Delta^d)$,
\begin{align}
   \ell^{S,0}_j(v) &= (\grad_S\phi^S_j, v)_{S}, \\
   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S}, \label{eq:hcurl-interior-dofs}
\end{align}
where $\curl_S$ is the tangential curl on $S$,
$\{\phi^S_j\}$ are the $H_0(\grad, S)$ bases constructed in \eqref{eq:hgrad-fdm},
and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S \mathbb{B}_{p_S}(\curl, S)$
such that
\begin{equation} \label{eq:hcurl-fdm}
   (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}, \quad
   \Phi^S_j\times\bn = 0 \text{ on }\partial S.
\end{equation}

@@

We define a basis for the dual of $V^2_{p, p_F}(K)$ as normal moments on $F\in
\text{faces}(\Delta^d)$,
\begin{equation}
   \ell^F_j(v) = (q_j, v\cdot \bn)_F, \quad q_j \in \P_{p_F}(F),
\end{equation}
and on the $\text{interior}(\Delta^d) = K$: 
\begin{align}
   \ell^{K,0}_j(v) &= (\curl\Phi^K_j, v)_{K}, \\
   \ell^{K,1}_j(v) &= (\div\Psi^K_j, \div v)_{K}, \label{eq:hdiv-interior-dofs}
\end{align}
where $\{\Phi^K_j\}$ are the $H_0(\curl, S)$ bases constructed in
\eqref{eq:hcurl-fdm}, and $\{\div\Psi^K_j\}$ is a basis for $\div
\mathbb{B}_p(\div, K)$ such that
\begin{equation} \label{eq:hdiv-fdm}
   (\div\Psi^K_j, \div\Psi^K_i)_K = \delta_{ij}, \quad
   (\Psi^K_j, \Psi^K_i)_{K} = \lambda_j\delta_{ij}, \quad
   \Psi^K_j\cdot\bn = 0 \text{ on }\partial K.
\end{equation}

>>>>

% Classic symbols
\newcommand{\reals}{\mathbb{R}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\dx}{\,\d\mathbf{x}}
\newcommand{\md}{\,\d}
\newcommand{\bigo}[1]{\mathcal{O}(#1)}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\union}{\cup}

@@

\newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: #1}}

@@

\begin{figure}[htbp] 
\centering
\begin{tikzcd}
  \CG_p \arrow[r, "\grad"] & \Ned_p \arrow[r, "\curl"] & 
   \RT_p \arrow[r, "\div"] & \DG_{p-1} \\
  \Hgrad \arrow[r, "\grad"] \arrow[u] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
  \arrow[d] \arrow[u] & \Hdiv \arrow[r, "\div"] \arrow[u] \arrow[d] & \Ltwo \arrow[u] \arrow[d] \\
   \CG_p \arrow[r, "\grad"] & \NedTwo_{p-1} \arrow[r, "\curl"] & 
   \BDM_{p-2} \arrow[r, "\div"] & \DG_{p-3}
\end{tikzcd}
\caption{The $\Ltwo$ de Rham complex (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
\end{figure}
\pef{Fix vertical alignment}

For each finite element subcomplex built on a mesh $\mesh$ of a domain $\Omega \subset \mathbb{R}^d$, we denote the spaces discretizing $k$-forms $H\Lambda^k(\Omega)$ with polynomial degree $p$ as $X^k_p(\mesh)$.
These are built from cellwise spaces via the standard pullback construction from a reference cell $\Khat$:
\begin{equation}
X^k_p(\mesh) = \{ v \in H\Lambda^k(\Omega): \Forall K \in \mesh \Exists \hat{v} \in X^k_p(\Khat) \text{ s.t.} \left.v\right|_K = \pullback^k_K(\hat{v}) \},
\end{equation}
where $\pullback^k_K$ denotes the appropriate pullback that preserves continuity of the traces of a $k$-form across cell facets.

We follow the definition of interpolation degrees of freedom from
Demkowicz et al.~\cite{demkowicz00} to construct dual bases that promote orthogonality in the
$\Ltwo(\Khat)$ and $H(\mathrm{d}, \Khat)$ inner products.
In the construction of the degrees of freedom we will employ bubble subspaces on a given mesh entity $S \subset \Khat$ with zero trace, defined as
\begin{equation}
B^k_p(S) \coloneqq \{v \in X^k_p(S): \tr v = 0 \text{ on } \partial S\}.
\end{equation}

Throughout we choose $\Khat$ to be an equilateral simplex, as this ...

@@

On $\Khat$ we discretize $\Hgrad$ with the space
\begin{equation}
X^0_{p}(\Khat) = \P_p(\Khat),
\end{equation}
where $\P_p(\Khat)$ is the space of polynomials of total degree less than or equal to $p$.
We define a basis for the dual of $X^0_p(\Khat)$ as
point evaluations at $V \in \text{vertices}(\Khat)$,
\begin{equation}
   \ell^V(v) = v(V),
\end{equation}
and for each sub-entity $S \in \text{edges}(\Khat) \union \text{faces}(\Khat) \union \text{interior}(\Khat)$
we take integral moments of surface gradients, 
\begin{equation}
   \ell^S_j(v) = (\grad_S\phi^S_j, \grad_S v)_{S},
\end{equation}
where $\grad_S$ is the tangential gradient on $S$,
and $\{\phi^S_j\}$ is a basis for $B^0_p(S)$
such that
\begin{equation} \label{eq:hgrad-fdm}
   (\grad_S\phi^S_j, \grad_S\phi^S_i)_{S} = \delta_{ij}, \quad
   (\phi^S_j, \phi^S_i)_{S} = \lambda_j\delta_{ij}.
\end{equation}
The eigenbases $\{\phi^S_j\}$ are numerically computed offline and only once 
on the reference interval, triangle, and tetrahedron.

@@

We define a basis for the dual of $V^1_{p, p_E}(K)$ as 
tangential moments along $E\in \text{edges}(\Khat)$,
\begin{equation}
   \ell^E_j(v) = (q_j, v\cdot \bt)_E, \quad q_j \in \P_{p_E}(E), 
\end{equation}
and for each sub-entity $S \in \text{faces}(\Khat) \cup \text{interior}(\Khat)$,
\begin{align}
   \ell^{S,0}_j(v) &= (\grad_S\phi^S_j, v)_{S}, \\
   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S}, \label{eq:hcurl-interior-dofs}
\end{align}
where $\curl_S$ is the tangential curl on $S$,
$\{\phi^S_j\}$ are the $H_0(\grad, S)$ bases constructed in \eqref{eq:hgrad-fdm},
and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S \mathbb{B}_{p_S}(\curl, S)$
such that
\begin{equation} \label{eq:hcurl-fdm}
   (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}, \quad
   \Phi^S_j\times\bn = 0 \text{ on }\partial S.
\end{equation}

@@

We define a basis for the dual of $V^2_{p, p_F}(K)$ as normal moments on $F\in
\text{faces}(\Khat)$,
\begin{equation}
   \ell^F_j(v) = (q_j, v\cdot \bn)_F, \quad q_j \in \P_{p_F}(F),
\end{equation}
and on the $\text{interior}(\Khat) = K$: 
\begin{align}
   \ell^{K,0}_j(v) &= (\curl\Phi^K_j, v)_{K}, \\
   \ell^{K,1}_j(v) &= (\div\Psi^K_j, \div v)_{K}, \label{eq:hdiv-interior-dofs}
\end{align}
where $\{\Phi^K_j\}$ are the $H_0(\curl, S)$ bases constructed in
\eqref{eq:hcurl-fdm}, and $\{\div\Psi^K_j\}$ is a basis for $\div
\mathbb{B}_p(\div, K)$ such that
\begin{equation} \label{eq:hdiv-fdm}
   (\div\Psi^K_j, \div\Psi^K_i)_K = \delta_{ij}, \quad
   (\Psi^K_j, \Psi^K_i)_{K} = \lambda_j\delta_{ij}, \quad
   \Psi^K_j\cdot\bn = 0 \text{ on }\partial K.
\end{equation}
